Draft

### 2017-08-13 Thesis/Draft.Rmd
Add experiment results discussion to thesis draft

$$
-\log(LR(\mu_0||\mu_1)) = n(\log\Big(\frac{1-\mu_0}{1-\mu_1}\Big)) - \Big(\log(\frac{\mu_0}{1-\mu_0}) - \log(\frac{\mu_1}{1-\mu_1})\Big)\sum_{i=1}^{n}x_i = (1-\hat{\mu})\log\Big(\frac{1-\mu_1}{1-\mu_0}\Big) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0})
$$

@@

$$
-\log(LR(\tau||\hat{\mu})) = KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = d(\hat{\mu},\tau)
$$

@@

$$
-\log(LR(\mu_0||\mu_1)) = n(\mu_0 - \mu_1) - (\log(\mu_0) - \log(\mu_1))\sum_{i=1}^n x_i = (\mu_0 - \mu_1) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}\Big)
$$

@@

$$
-\log(LR(\tau||\hat{\mu})) = KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = d(\hat{\mu},\tau) = (\tau - \hat{\mu}) + \hat{\mu}\log\Big(\frac{\hat{\mu}}{\tau}\Big)
$$

@@

$$
-\log(LR(\mu_0||\mu_1)) = n (\frac{\mu_0^2}{2\sigma^2} - \frac{\mu_1^2}{2\sigma^2}) - (\frac{\mu_0}{\sigma^2} - \frac{\mu_1}{\sigma^2})\sum_{i=1}^nx_i = \frac{1}{2\sigma^2}\Big(\mu_0^2-\mu_1^2 - 2\mu_0\hat{\mu} + 2\mu_1\hat{\mu}\Big)
$$

@@

$$
-\log(LR(\tau||\hat{\mu})) = KL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = d(\hat{\mu},\tau) = \frac{(\hat{\mu} - \tau)^2}{2\sigma^2}.
$$

@@

knitr::kable(rbind(Training, Test, TrainingMinusTau), digits = 3, col.names = names(data_amo3_mean_firsthalf),
             caption = "Count rates for ten products for both training and test sample.")

@@

@@

>>>>

$$
-\log(LR(\mu_0||\mu_1)) = n(\log\Big(\frac{1-\mu_0}{1-\mu_1}\Big)) - \Big(\log(\frac{\mu_0}{1-\mu_0}) - \log(\frac{\mu_1}{1-\mu_1})\Big)\sum_{i=1}^{n}x_i = n \Big( (1-\hat{\mu})\log\Big(\frac{1-\mu_1}{1-\mu_0}\Big) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}) \Big)
$$

@@

$$
-\log(LR(\tau||\hat{\mu})) = nKL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = nd(\hat{\mu},\tau) = n \Big(\hat{\mu}\log\big(\frac{\hat{\mu}}{\tau}\big) + (1-\hat{\mu})\log\big(\frac{1-\hat{\mu}}{1-\tau}\big)\Big)
$$

@@

$$
-\log(LR(\mu_0||\mu_1)) = n(\mu_0 - \mu_1) - (\log(\mu_0) - \log(\mu_1))\sum_{i=1}^n x_i = n\Big((\mu_0 - \mu_1) + \hat{\mu}\log\Big(\frac{\mu_1}{\mu_0}\Big)\Big)
$$

@@

$$
-\log(LR(\tau||\hat{\mu})) = nKL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = nd(\hat{\mu},\tau) = n\Big((\tau - \hat{\mu}) + \hat{\mu}\log\Big(\frac{\hat{\mu}}{\tau}\Big)\Big)
$$

@@

$$
-\log(LR(\mu_0||\mu_1)) = n (\frac{\mu_0^2}{2\sigma^2} - \frac{\mu_1^2}{2\sigma^2}) - (\frac{\mu_0}{\sigma^2} - \frac{\mu_1}{\sigma^2})\sum_{i=1}^nx_i = n\frac{1}{2\sigma^2}\Big(\mu_0^2-\mu_1^2 - 2\mu_0\hat{\mu} + 2\mu_1\hat{\mu}\Big)
$$

@@

$$
-\log(LR(\tau||\hat{\mu})) = nKL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = nd(\hat{\mu},\tau) = n\frac{(\hat{\mu} - \tau)^2}{2\sigma^2}.
$$

#### Example 4: Gamma distributions with known shape parameter $\alpha$

We have $T(x_i) = x_i$, $\theta = -\alpha/\mu$, $\Theta = (-\infty, 0)$, $b(\theta) = -\alpha \log(-\theta)$, $\mu \in (0,+\infty)$. Then $b'(\theta) = -\mu$, and we can write

$$
-\log(LR(\mu_0||\mu_1)) = n\alpha(-\log(\frac{\alpha}{\mu_0}) - \log(\frac{\mu_1}{\alpha})) - (-\frac{\alpha}{\mu_0} + \frac{\alpha}{\mu_1})\sum_{i=1}^{n}x_i = n\alpha \Big(\log\big(\frac{\mu_0}{\mu_1}\big) - \frac{\hat{\mu}}{\mu_1} + \frac{\hat{\mu}}{\mu_0}\Big).
$$
And again, if we let $\mu_1 = \hat{\mu}$ and $\mu_0 = \tau$, then

$$
-\log(LR(\tau||\hat{\mu})) = nKL(\nu_{b'^{-1}(\hat{\mu})}||\nu_{b'^{-1}(\tau)}) = nd(\hat{\mu},\tau) = n\alpha \Big(\frac{\hat{\mu}}{\tau} - 1 - \log\big(\frac{\hat{\mu}}{\tau}\big)\Big).
$$

For $\alpha = 1$, the Gamma distribution is equal to the Exponential distribution.

Given these results for the exponential family of distributions, we can now define the SLR (Simple Likelihood Ratio) algorithm.

### 1D-SLR Algorithm for One-Dimensional Exponential Family of Distributions

Very similar to the APT algorithm, the SLR algorithm receives as input the definition of the problem $(\tau, \epsilon)$ (compare Locatelli et al., 2016). Additionally, the algorithm receives as input the name of the distribution that the arms follow, as well as potential additional parameters ($\sigma_1, ..., \sigma_K$ for Normal distributions, $\alpha_1, ..., \alpha_K$ for Gamma distributions). We assume that each arm follows the same kind of distribution, with potentially different parameters $\theta_1, ..., \theta_K$. The algorithm pulls each arm once. At time $t>K$, the algorithm updates the number of pulls for arm $i$ at time $t$, $T_i(t)$, for each arm, as well as the empirical mean $\hat{\mu}_i(t)$ of arm $i$ after $T_i(t)$ pulls. The empirical mean is calculated for each arm $k\in \{1,...,K\}$ as $\hat{\mu}_i(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)} x_{i,s}$, with $T_i(t) = \sum_{s=1}^{t}\mathbb{I}(I_s = i)$. As in Locatelli et al. (2016), the $X_{i,s}$ is the sample received when pulling arm $i$ for the $s$-th time. Depending on the distribution, the algorithm then computes the index

$$
B_i^{SLR}(t+1) = \mathbb{I}(\hat{\mu}\geq \tau)(-\log\big(LR(\tau-\epsilon||\hat{\mu}_i(t))\big) + \mathbb{I}(\hat{\mu}< \tau)(-\log\big(LR(\tau+\epsilon||\hat{\mu}_i(t))\big) = T_i(t) \Big( \mathbb{I}(\hat{\mu}\geq \tau)\cdot KL(\nu_{b'^{-1}(\hat{\mu}(t))}||\nu_{b'^{-1}(\tau-\epsilon)}) + \mathbb{I}(\hat{\mu} < \tau)\cdot KL(\nu_{b'^{-1}(\hat{\mu}(t))}||\nu_{b'^{-1}(\tau+\epsilon)}) \Big).
$$

The next arm to be pulled is the minimizer of the index $B_i^{SLR}(t+1)$, $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$. At the end of the horizon $T$, the algorithm outputs the set of arms $\hat{S}_\tau = \{k: \hat{\mu}_k(T) \geq \tau\}$.

### 2D-SLR Algorithm for Normal Distributions with Unknown Mean and Unknown Variance

As soon as we move away from univariate distributions parameterized by their mean, it becomes less clear how the likelihood ratio for $H_0: \mu = \tau$ and $H_1: \mu = \hat{\mu}$ should be calculated. In the case of Normal distributions with unknown mean and unknown variance, given that the variance and mean estimate are linked, we can use $H_0: \mu = \tau \land \sigma^2 = \sigma^2_{\tau} = \frac{1}{n}\sum_{i=1}^{n}(x_i-\tau)^2$ and the alternative hypothesis based on the maximum likelihood estimates $H_1: \mu = \hat{\mu} \land \sigma^2 = \hat{\sigma}^2 = \frac{1}{n}\sum_{i=1}^{n}(x_i-\hat{\mu})^2$.

In the following, we assume $\epsilon = 0$.

Now, at each round $t$, the algorithm computes $T_i(t)$ and $\hat{\mu}_i(t)$ as before, but additionally the empirical variances $\sigma_{\tau,i}^2(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)}(x_{i,s}-\tau)^2$ and $\hat{\sigma}_{i}^2(t) = \frac{1}{T_i(t)}\sum_{s=1}^{T_i(t)}(x_{i,s}-\hat{\mu})^2$. The index is then given by

$$
B_i^{2DSLR}(t+1) = -\log(LR(\tau, \sigma_{\tau,i}^2(t)||\hat{\mu}_i(t),\hat{\sigma}^2_{i}(t))) = T_i(t) \log\Big(\frac{\sigma_{\tau,i}^2(t)}{\hat{\sigma}_i^2(t)}\Big) = T_i(t) KL(\hat{\mu}_i(t),\hat{\sigma}^2_{i}(t)||\tau, \sigma_{\tau,i}^2(t))
$$

(SHOW THAT THE NOTATION CAN BE ABUSED AND THE KL WRITTEN THIS WAY BECAUSE UNIQUELY DEFINED BY MEAN AND VARIANCE)

Then, as for the univariate case, the next arm to be pulled is the minimizer of the index $B_i^{2DSLR}(t+1)$, $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$. At the end of the horizon $T$, the algorithm outputs the set of arms $\hat{S}_\tau = \{k: \hat{\mu}_k(T) \geq \tau\}$.

@@

knitr::kable(rbind(Training, Test, TrainingMinusTau)[,order(Training)], digits = 3, col.names = names(data_amo3_mean_firsthalf),
             caption = "Count rates for ten products for both training and test sample.")

@@

These difficulties seem to have a large impact on the performance of all algorithms. Given that the data are $\{0,1\}$ observations, we can compare all algorithms in a setting in which one expects all assumptions to hold before having observed the data. The distribution is sub-Gaussian for the APT algorithm, bounded between $0$ and $1$ for the variance-based algorithms, and assumed to be Bernoulli for the SLR algorithm. Additionally, we use the B-UCB algorithm as it performed often best on the synthetic data.

We set $\epsilon = 0$, and $\tau = 2/60$. For the AugUCB algorithm we use the parameters as in Mukherjee et al. (2017), that is, $\rho = 1/3$. For B-UCB we set the prior to $\alpha = \tau$ and $\beta = 1-\tau$. All other algorithms are parameter free.

The corresponding results are shown in the next figure.

```{r, echo = FALSE}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/"

#load(paste0(current_path, "amo3_comp_BUCB.Rda"))
load(paste0(current_path, "amo3_comp_APT.Rda"))
load(paste0(current_path, "amo3_comp_UNIFORM.Rda"))
load(paste0(current_path, "amo3_comp_AugUCB.Rda"))
load(paste0(current_path, "amo3_comp_LR.Rda"))
load(paste0(current_path, "amo3_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB","LR"), 
                   each = 10071), rep("EVT", 10061))
index <- c(rep(10:10080, times = 4), 20:10080)
alg_res <- log(c(amo3_comp_UNIFORM, amo3_comp_APT, amo3_comp_AugUCB, amo3_comp_LR, amo3_comp_EVT))
loc3_comp_res <- data.frame(index = index, names = alg_names, 
                            results = alg_res)
ggplot(loc3_comp_res, aes(x = index, y = results)) + 
  geom_line(aes(group = names, color = names)) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(title = "Results of Real Data Experiment 1", 
       subtitle = paste0("Sample Mean Complexity of "), 
       y = "log(Error)", x = "Round Index")
```

Compared to the previous simulations with synthetic data, the results are dramatic. First of all, we see that all strategies perform not well. All of them end above the 10% error mark after 1 week of data. Second, the uniform sampling strategy fares best when evaluated against the sample mean of all observations used during training. This of course is the opposite of what we would like to see, and the opposite of what we saw on the synthetic data, where we were able to improve upon the naive strategy. Third, again in contrast to the results on synthetic data, we observe non-monotonicity in the error curves. Given that the highs and lows run very much parallel for the different models, it stands to reason that this effect is caused by the variation in means of the arms over time. Indeed, roughly comparing the error curves to the moving averages in the previous figure, the increase in errors after round $5000$ and the decreasing error after round $7500$ seems to line up more or less with the fluctuations observed in the data. Given that we evaluate the models against a sample mean of all observations, it is not surprising that the uniform sampling strategy has a certain advantage when it comes to "factoring out" the seasonalities. In contrast, the adaptive sampling schemes try to adapt more or less quickly to the changes in the means; a feature that would be valuable in a cumulative regret multi-armed bandit setting. Here, however, it leads to wrong estimates of the sample mean.

If we focus on a comparison of adaptive sampling strategies and less on their globally bad performance, we see that the variance based strategies perform worse than APT and SLR. Very dramatic is the behavior of the Augmented-UCB algorithm, which first seems to perform well; that, however, is a result of it mostly pulling arms uniformly, until it starts to adapt its sampling and discarding arms after nearly 5000 rounds. The result is a dramatic increase in error which leads to the worst performs at the end of the budget. The difference between APT, SLR, and EVT is not very large, and their performance fluctuates in sync. None of them is really able to converge due to the changes in the underlying means. Advantageous only seems to be the fact that their behavior is less erratic than that of the Augmented-UCB strategy given that they do not discard arms.

@@

In order to underline the results of the previous experiment, we perform a second experiment using the data collected by Amorelie. We increase the threshold to $\tau = \frac{5}{60} = 0.833$ to make the problem in general slightly easier, and again pick 10 arms at random from the data set of 197 arms. We again use a rolling window of $10080$ observations to create $5000$ iterations and look at the training and test sample means as in Experiment 1.

Some arms, in particular $V110$ and $V150$, but also $V103$, $V149$, $V104$, and $V112$ are quite close to the threshold. Two arms, $V129$ and $V32$ appear to have basically no positive observation in the training set. On the other hand, arm $V154$ is larger than the threshold by a fair amount and should be easier to classify.

```{r experiment2_arm_means, echo = FALSE, message = FALSE}

load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_mean_firsthalf.Rda")
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_mean_secondhalf.Rda")

Training <- data_amo4_mean_firsthalf
Test <- data_amo4_mean_secondhalf
tau <- 5/60
TrainingMinusTau <- Training - tau

knitr::kable(rbind(Training, Test, TrainingMinusTau)[,order(Training)], digits = 3, col.names = names(data_amo4_mean_firsthalf),
             caption = "Count rates for ten products for both training and test sample.")

As in the previous experiment, we can also take a look at the one day moving averages of the arms to quickly assess the time dependency of the means. Given that the data here are collected at the same time points as the data used in the previous experiment, it is not surprising to see a similar spike in means during the second half of the observations with the maximum moving average around index 7500. We also observe that these fluctuations are heavier for products with in general larger means (that is, more visitors on the website). But these fluctuations are clearly much larger and systematic than those observed for synthetic data with the same sample means. While the moving average of product $V159$ in the synthetic data has a valley after index 5000, this can be explained by the natural variance (which increases as the mean is closer to $0.5$).

Given the moving averages and the previous experiment, the performance results of Experiment 2 shown in the next figure are not surprising. We recognize the same structure as previously. The uniform sampling strategy performs the best. The adaptive sampling strategies perform bad, do not converge, and end above the 10% error mark. This time, Augmented-UCB is not (yet) the worst performing strategy, but equally erratic after it actually adapts its sampling. This time around, it even has two spikes in its adaptive phase. Again, APT, SLR, and EVT perform in sync and more or less equally bad, with EVT slightly better than APT and SLR this time around. Interestingly, Bayes-UCB performs clearly worst of all strategies. Its performance changes at the same time as Augmented-UCB starts to adapt its sampling strategy, near round 7000.

This last fact is of particular interest, as Bayes-UCB is dramatically better than the other strategies in experiments on synthetic data. On the right hand side of the same figure, the performance of the algorithms on synthetic data from Bernoulli distributions with the same sample mean is depicted. On this data, Bayes-UCB performs very well, ending at an average error of about 0.1% of the simulations. This mirrors its performance observed in the previous simulations with synthetic data. Also all other strategies perform clearly better now, and their error rates are strictly decreasing. Interestingly, the uniform sampling scheme is now much worse than in the experiment on real data; it does no longer have the advantage of sampling the different mean phases at a representative rate. Instead, it is no in its original position of serving as upper bound on the error rate. Excluding Bayes-UCB, we see that the SLR algorithm performs better the best, followed by the variance-based algorithms. APT comes in at the same error rate as uniform sampling. While the difference between SLR and EVT are less than 2% (1.2% vs 3%), the difference to uniform sampling (10%) is relevant. The performance of 10% error rate is achieved by SLR about 1000 observations earlier than by EVT.

```{r, echo = FALSE, message = FALSE}
library(zoo)
library(tidyr)
library(dplyr)
library(ggplot2)
load("/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/data_amo4_wide.Rda")

real_daily <- data.frame(index = 1:8641,
                                lapply(data_amo4_wide[1:10080,], FUN = rollapply, 
                                       width = 24*60, mean)) %>%
  gather(arm, mean, -index) %>%
  mutate(source = rep("Real", 86410))

real_means <- colMeans(data_amo4_wide[1:10080,])
tau_amo4 <- 5/60

syn_data <- data.frame(rep(NA, times = 10080))
set.seed(124632728)
for(i in 1:length(real_means)) {
  syn_data[[i]] <- as.numeric(purrr::rbernoulli(10080, p  = real_means[i]))
}
names(syn_data) <- names(data_amo4_wide)

rbind(real_daily, syn_daily) %>%
  ggplot(aes(x = index, y = mean, group = arm, color = arm)) +
  geom_line(size = 0.4) +
  geom_hline(aes(yintercept = tau_amo4), linetype = 2) +
  facet_grid(.~source) + 
  labs(title = "Moving average over 24*60 observations (1 day)", 
       subtitle = "Horizontal bar indicates the threshold.", 
       x = "Index", y = "One day moving average")
```

```{r, echo = FALSE}
library(ggplot2)
current_path <- "/Users/timradtke/Dropbox/1Master/Master Thesis/threshold-thesis/real_experiments/"

load(paste0(current_path, "amo4_comp_BUCB.Rda"))
load(paste0(current_path, "amo4_comp_APT.Rda"))
load(paste0(current_path, "amo4_comp_UNIFORM.Rda"))
load(paste0(current_path, "amo4_comp_AugUCB.Rda"))
load(paste0(current_path, "amo4_comp_LR.Rda"))
load(paste0(current_path, "amo4_comp_EVT.Rda"))

load(paste0(current_path, "amo4sim_comp_BUCB.Rda"))
load(paste0(current_path, "amo4sim_comp_APT.Rda"))
load(paste0(current_path, "amo4sim_comp_UNIFORM.Rda"))
load(paste0(current_path, "amo4sim_comp_AugUCB.Rda"))
load(paste0(current_path, "amo4sim_comp_LR.Rda"))
load(paste0(current_path, "amo4sim_comp_EVT.Rda"))

alg_names <- c(rep(c("Uniform", "APT", "AugUCB","LR", "B-UCB"), 
                   each = 10071), rep("EVT", 10061))
index <- c(rep(10:10080, times = 5), 20:10080)
alg_res <- log(c(amo4_comp_UNIFORM, amo4_comp_APT, amo4_comp_AugUCB, amo4_comp_LR, amo4_comp_BUCB, amo4_comp_EVT))
source_name <- rep("Real", 60416)
loc4_comp_res <- data.frame(source = source_name, 
                            index = index, names = alg_names, 
                            results = alg_res)

alg_names <- c(rep(c("Uniform", "APT", "AugUCB", "LR", "B-UCB"), 
                   each = 10071), rep("EVT", 10061))
index <- c(rep(10:10080, times = 5), 20:10080)
alg_res <- log(c(amo4sim_comp_UNIFORM, amo4sim_comp_APT, amo4sim_comp_AugUCB,
                 amo4sim_comp_LR, amo4sim_comp_BUCB, amo4sim_comp_EVT))
source_name <- rep("Synthetic", 60416)
loc4sim_comp_res <- data.frame(source = source_name, index = index, 
                               names = alg_names, results = alg_res)
loc4total <- rbind(loc4_comp_res, loc4sim_comp_res)

ggplot(loc4total, aes(x = index, y = results, 
                      group = names, color = names)) + 
  geom_line() +
  facet_grid(.~source) +
  geom_hline(aes(yintercept = log(0.1)), linetype = 2) +
  labs(title = "Results of Real Data Experiment 2", 
       subtitle = paste0("Sample Mean Complexity of "), 
       y = "log(Error)", x = "Round Index")
```
